### 2025-05-06 monograph_Luis/Abstract/abstract.tex

>>>>

This work presents the calibration of the CMS Pixel Cluster Counting (PCC) luminometer and the absolute luminosity determination using the van der Meer (vdM) method, as part of the publication \textit{``Precision luminosity measurement in proton-proton collisions at $\sqrt{s} = 13$ TeV with the CMS detector''}, which covers the second part of Run 2 corresponding to the 2017 and 2018 data-taking periods. The final results for the PCC visible cross section $\sigma_{\text{vis}}$ are 4666.4 $\pm$ 14.7 mb for 2017 and 964.9 $\pm$ 2.2 mb for 2018. The final integrated luminosity is obtained from a combination of multiple independently calibrated luminometers. A precision of 0.86\% for 2017 and 0.83\% for 2018 is achieved, with dominant uncertainties arising from luminometer non-linearity and the x–y factorization correction.

### 2025-05-06 monograph_Luis/Chapter1/chapter1.tex

\section{Particle Colliders and production cross section}

Particle colliders are powerful scientific instruments used to produce, accelerate, and collide beams of particles at relativistic speeds, enabling scientists to explore the fundamental properties of matter and the universe. These experiments function as powerful microscopes that allow researchers to study subatomic particles such as electrons, protons, neutrons, neutrinos, and quarks. They also provide a means to test predictions made by the Standard Model.

The high speeds required for the collisions are achieved through the use of powerful superconducting magnets, which accelerate particles to the necessary velocities. These particles are then collided at specific points along the accelerator, where detectors are located. These detectors are capable of measuring individual interactions, known as events, generated by the collisions. Some of the particles produced, such as the Higgs boson, cannot be directly observed. Instead, they are detected and analyzed using advanced technologies, which measure their interactions with matter \cite{thomson_2013}.\\

In general, a particle detector is typically constructed as a cylindrical or polygonal barrel section, oriented parallel to the incoming colliding beams. Two flat end caps seal the cylindrical structure, ensuring nearly complete solid angle coverage all the way down to the beam pipe. This detector comprises multiple layers, each designed to detect the different particles generated during collisions \cite{thomson_2013}.\\

Particle colliders are evaluated based on two important measures, which are the available center of mass energy and the rate at which particles collide (luminosity). The amount of energy that is available to create new particles and study their properties is particularly significant because it determines the range of particles that can be produced. During experiments, particles are accelerated and directed toward each other, and when they collide, they produce new particles and interactions that can be analyzed by scientists.\\

Colliding beam machines are capable of achieving significantly higher energies at the point where the beams intersect and scatter. When the particles in the beams have equal energy and opposite momentum (equal mass), the experiment occurs in the center-of-mass (CM) frame. This allows for the entire energy delivered by the accelerator to be utilized in generating high-mass particles \cite{undergraduate_accelerators_chapter}. In particle physics, the Lorentz invariant quantity $s$ represents the square of the total incoming energy and is defined as:

\begin{equation}
s = \left ( \sum_{i = 1,2}^{}E_{i} \right )^{2}-\left ( \sum_{i = 1,2}^{}p_{i} \right )^{2}c^{2}
\end{equation}
Where  \textbf{$p$} and $E$ are the momenta and the energy of each incoming particle respectively.In the center-of-mass (CM) frame, where the momenta of the particles are equal and opposite, the second term in the equation for the Lorentz invariant quantity ($s$) disappears, leaving only $s = 4E^2$.\\

@@

The technical meaning of "cross section" in particle physics is quite different from its common usage, while "cross section" typically refers to a slice of an object, in particle physics, it quantifies the probability of a collision and is a property of the interaction of one particle with another, this is expressed in units of square meters. When two beams cross in a particle accelerator, a variety of different processes can occur and  the cross section of a given process depends on the type and energy of the colliding particles.
At the Large Hadron Collider (LHC), certain particles such as  $W^{\pm}$ and $Z^{0}$ bosons have large cross sections , making their observation more frequent, while the production of a Higgs boson has a much lower cross section, making it more difficult to produce \cite{thomson_2013}.\\

To explain this concept more clearly, let's consider a simple scenario where a beam of particles of type $a$ passes through a region of space that contains particles of type $b$. The interaction rate between the two types of particles depends on the number of particles per unit volume of type $b$, denoted as $n_{b}$, and the flux of the incident particles, denoted as $\phi_{a}$. The rate of interaction per target particle of type $b$ can be expressed as follows:

@@

The essence of fundamental physics can be encapsulated in the interaction cross section, denoted by $\sigma$. This parameter, which is measured in area units, represents the likelihood that an interaction will occur between an incident particle and a target particle. It's helpful to imagine each target particle as having an effective cross-sectional area, and the interaction cross section as representing that area. Figure \ref{cross_section}(a) shows an illustration of this concept, where an incident particle of type $a$ moves with velocity $v_{a}$ through a region of area $A$ that contains target particles of type $b$ with velocity $v_{b}$ moving in the opposite direction. The value of the interaction cross section is determined by the underlying quantum mechanical probability of interaction and is fundamental to understanding the behavior of particles in collisions. \cite{thomson_2013}

@@

To understand the probability of an interaction between particles of types $a$ and $b$, imagine that in a time interval $\delta t$, particle $a$ passes through a region containing $\delta N= nb(v_{a}+v_{b}) A \delta t$ particles of type $b$. The probability of interaction can be obtained by dividing the effective total cross-sectional area of the $\delta N$ particles by the area $A$. This can be thought of as the probability that the incident particle passes through one of the regions of area $\sigma$ drawn around each of the $\delta N$ target particles, as shown in Figure \ref{cross_section}(b). So the cross section for a particular process is defined as the effective target area for that process and has units of area, more formally is expresed as \ref{cross} \cite{thomson_2013}.

%For a beam of particles of type $a$ with number density $n_{a}$ confined to a volume $V$, the total interaction rate can be expressed as $\text{rate}=(n_{a}v)(n_{b}V) \sigma = \phi N_{b} \sigma$, where $v$ is the velocity of the beam particles, $N_{b}$ is the number of target particles per unit volume, and $\phi$ is the flux of incident particles per unit area per unit time. The cross section for a particular process is defined as the effective target area for that process and has units of area, more formally is expresed as \cite{thomson_2013}:
\begin{equation}
\sigma=\frac{\text{number of interaction per unit time per target particle}}{\text{incident flux}}
\label{cross}
\end{equation}

\section{Luminosity }

In experiments conducted in particle physics, the number of useful interactions (or events) is just as important as the energy. The ability of a particle accelerator to produce these interactions is measured by a parameter called instantaneous luminosity $\mathcal{L}_{inst}$. It is directly proportional to the number of events produced per second and the cross section $\sigma_{p}$ \cite{concept_of_luminosity}.

\begin{equation}
R=\mathcal{L}_{inst} \cdot \sigma_{p}
\end{equation}

Where the unit of luminosity is $cm^{-2}s^{-1}$.

To derive a general expression for luminosity in the case of two colliding beams, we consider both beams to serve as the target and the incoming beam simultaneously. We will assume that the beams are "bunched", meaning that the particles are grouped into packets or bunches.

\begin{center}
  \begin{figure}[h!]
    \centering
\includegraphics[scale=.25]{Chapter1/luminosity.png} 
 \caption[Colliding beam interaction]{Schematic view of a colliding beam interaction\cite{concept_of_luminosity}}
    \label{luminosity}
  \end{figure}
\end{center}

%A schematic picture is shown in Fig. \ref{luminosity}. Since the two beams are not stationary but are moving through each other, the overlap integral depends on the longitudinal position of the bunches, and therefore on time. As the beams move towards and through each other, the two beams have different distribution functions and a different number of particles in the beams \cite{concept_of_luminosity}. The overlap integral, which is proportional to the luminosity $\mathcal{L}_{inst}$, can then be written as:
 In Figure \ref{luminosity}, you can see a diagram that illustrates how two beams intersect with each other. Because these beams are not still, but rather in motion, the way they overlap depends on the time and the position of the bunches. While the beams move, they also have different distributions and numbers of particles in them \cite{concept_of_luminosity}. The luminosity $\mathcal{L}_{inst}$, which is proportional to the overlap integral, can be expressed as:

\begin{equation}
  \mathcal{L}_{inst}\propto K\cdot \int\int\int \int_{-\infty}^{\infty} \rho_{1}(x,y,s,-s_{0}) \rho_{2}(x,y,s,s_{0})dxdydsds_{0}
    \label{lumi_1}
\end{equation}

The functions $\rho_{1}(x,y,s,-s_{0})$ and $\rho_{2}(x,y,s,s_{0})$ represent the beam density distribution at a given time. We assume that the two bunches meet at the point $s_{0} = 0$. Since the beams are in motion towards each other, we need to take into account a relativistic kinematic factor (Eq. \ref{kinematic}) by multiplying this expression.

\begin{equation}
 K = \sqrt{(\vec{v_{1}}-\vec{v_{2}})^{2}-(\vec{v_{1}}\times \vec{v_{2}})^{2}/c^{2}}
    \label{kinematic}
\end{equation}

When the bunches collide head-on and are traveling at nearly the speed of light, the relativistic kinematic factor takes on a value of 2.\\
Assuming uncorrelated densities in all planes and head-on collisions where the velocities of the two bunches are opposite $(\vec{v_{1}}=-\vec{v_{2}})$, the overlap integral can be factorized as:

\begin{equation}
  \mathcal{L}_{inst}= 2N_{b} N_{1}N_{2}f \int\int\int\int_{-\infty}^{\infty}  \rho_{1x}(x)\rho_{1y}(y)\rho_{1s}(s-s_{0})\rho_{2x}(x)\rho_{2y}(y)\rho_{2s}(s+s_{0}) dxdydsds_{0}
    \label{luminosity_2}
\end{equation}

where $N_{b}$ is the number of bunches in one beam, $N_{1}$ and $N_{2}$ are the particles per bunch and $f$ the revolution frequency.\\

%\begin{eqnarray}
%\rho_{iz}(z)= \frac{1}{\sqrt{2\pi} \sigma_{z}} exp \biggl(-\frac{z^{2}}{2\sigma_{z}^{2}} \biggr)&z=x,y&i=1,2 
%\end{eqnarray}

%\begin{eqnarray}
%\rho_{s}(s\pm s_{0})= \frac{1}{\sqrt{2\pi} \sigma_{s}} exp \biggl(-\frac{(s\pm s_{0})^{2}}{2\sigma_{s}^{2}} \biggr)
%\end{eqnarray}

Assuming Gaussian profiles in all dimensions and approximately equal bunch lengths $(\sigma_{1s}\approx \sigma_{2s})$, the equation \ref{luminosity_2} can be solved.
For a general case where $\sigma_{1x}\neq \sigma_{2x} \text{ and } \sigma_{1y}\neq \sigma_{2y}$

\begin{equation}
  \mathcal{L}_{inst}= \frac{N_{1} N_{2} N_{b}f }{2\pi \sqrt{\sigma_{1x}^{2}+\sigma_{2x}^{2}}\sqrt{\sigma_{2y}^{2}+\sigma_{2y}^{2}}}
  \label{lumi_general}
\end{equation}

For a specific case, asuming equal beams, $\sigma_{1x}= \sigma_{2x} \text{ and } \sigma_{1y}= \sigma_{2y}$:

\begin{equation}
  \mathcal{L}_{inst}= \frac{N_{1} N_{2} N_{b}f }{4\pi \sigma_{x} \sigma_{y}}
  \label{lumi_theor}
\end{equation}

This expression is widely recognized as the formula for the luminosity of two Gaussian beams that collide head-on. It reveals how the luminosity is related to the number of particles in each bunch and the size of the beams \cite{concept_of_luminosity}.\\

In practice,the integral in  \ref{luminosity_2} cannot be solved analytically because   the properties of the colliding beams are not know precisely, so a experimental technique is implemented with a dedicated machine setup to estimate the integrals, yielding a similar expression as \ref{lumi_theor}.

To determine the overall number of events, we rely on integrated luminosity, which is a measurement of the total amount of collected data. This value is crucial for evaluating the effectiveness of an accelerator and it can be expressed as: \cite{concept_of_luminosity}:

\begin{equation}
  \mathcal{L}_{int}=\int_{0}^{T} \mathcal {L}_{\text{inst}}(t') dt'
\end{equation}

Where the value of $\mathcal{L}_{inst}$ is evaluated over time $T$ (excluding possible dead time) because its intensity decreases as the number of protons reduces during the collisions of the fill. So $\mathcal{L}_{int}$ is directly related to the number of observed events as:
\begin{equation}
  \mathcal{L}_{int} \ndot \sigma_{p}= \text{number of events of interest}
\end{equation}

The integrated luminosity has units of $cm^{-2}$ and is often expressed in inverse barn ($1 barn= 10^{-24}cm^{2}$). \\

@@

\section{Results in Luminosity precision for LHC Run 2}

@@

The second stage (2017–2018) had preliminary results published in the Public Analysis Summaries, CMS-PAS-LUM-17-001 \cite{pas_17} for 2018 and CMS-PAS-LUM-18-002 \cite{pas_18} for 2017. However, the desired uncertainty values were not achieved.
Efforts are currently underway to further improve the precision of these measurements and present them in a final publication. This project is primarily focused on contributing to this publication by analyzing the PCC luminometer and refining the uncertainty, aiming to achieve values of approximately 1\%.\\

@@

\section{Motivation for precise luminosity measurement}

%% tabla con ano , luminosidad y presicion

>>>>

\section{Luminosity}

Particle colliders are powerful scientific instruments used to produce, accelerate, and collide beams of particles at relativistic speeds, enabling scientists to explore the fundamental properties of matter and the universe. These experiments function as powerful microscopes that allow researchers to study subatomic particles such as electrons, protons, neutrons, neutrinos, and quarks. They areevaluated based on two important measures, which are the available center of mass energy and the rate at which particles collide (luminosity).

When the particles in the beams have equal energy and opposite momentum (equal mass), the experiment occurs in the center-of-mass (CM) frame. This allows for the entire energy delivered by the accelerator to be utilized in generating high-mass particles \cite{undergraduate_accelerators_chapter}. In particle physics, the Lorentz invariant quantity $s$ represents the square of the total incoming energy and is defined as:

\begin{equation}
s = \left ( \sum_{i = 1,2}^{}E_{i} \right )^{2}-\left ( \sum_{i = 1,2}^{}p_{i} \right )^{2}c^{2}
\end{equation}

Where  \textbf{$p$} and $E$ are the momenta and the energy of each incoming particle respectively.In the center-of-mass (CM) frame, where the momenta of the particles are equal and opposite, the second term in the equation for the Lorentz invariant quantity ($s$) disappears, leaving only $s = 4E^2$.\\

@@

When two beams cross in a particle accelerator, a variety of different processes can occur and  the cross section of a given process depends on the type and energy of the colliding particles.
At the Large Hadron Collider (LHC), certain particles such as  $W^{\pm}$ and $Z^{0}$ bosons have large cross sections , making their observation more frequent, while the production of a Higgs boson has a much lower cross section, making it more difficult to produce \cite{thomson_2013}.\\

To illustrate this concept more clearly, let's consider a simple scenario where a beam of particles of type $a$ passes through a region of space that contains particles of type $b$. The interaction rate between the two types of particles depends on the number of particles per unit volume of type $b$, denoted as $n_{b}$, and the flux of the incident particles, denoted as $\phi_{a}$. The rate of interaction per target particle of type $b$ can be expressed as follows:

@@

Figure \ref{cross_section}(a) shows an illustration of this concept, where an incident particle of type $a$ moves with velocity $v_{a}$ through a region of area $A$ that contains target particles of type $b$ with velocity $v_{b}$ moving in the opposite direction \cite{thomson_2013}.

@@

In particle physics experiments, the number of useful interactions (or events) is a crucial, the capability of a particle accelerator to produce such interactions is quantified by a parameter known as the instantaneous luminosity, denoted as $\mathcal{L}_{\text{inst}}$. The significance of luminosity lies in its direct relationship with the production rate of particles of interest. Specifically, the rate at which a particular process occurs is proportional to the product of the luminosity and the corresponding cross section. As such, luminosity directly impacts the ability to observe rare events. Furthermore, achieving precise measurements of particle properties often requires a large number of collisions to ensure sufficient statistical significance.

The event rate $R$ for a given process is defined as the product of the instantaneous luminosity and the process cross section $\sigma_p$~\cite{concept_of_luminosity}:

\begin{equation}
R = \mathcal{L}_{\text{inst}} \cdot \sigma_p
\label{eq.lumi}
\end{equation}

Here, luminosity is expressed in units of $cm^{-2}s^{-1}$.

The cross section of the process that produces the particle is a measure of the probability that a given interaction will occur. It is typically a very small quantity, on the order of picobarns ($10^{-36}~\text{cm}^2$). To achieve high luminosity, particle accelerators employ various techniques to increase the number of particles in the beams and to focus them into a small cross-sectional area. This is accomplished through the use of superconducting magnets, which guide and confine the particles to a narrow path. Additionally, radio frequency (RF) cavities are used to accelerate the particles to higher energies, thereby increasing the frequency of bunch crossings and enhancing the overall collision rate.

In high-energy particle accelerators, such as the Large Hadron Collider (LHC) at CERN, integrated luminosity is a key quantity that measures the total number of collisions per unit area accumulated during an experiment. It quantifies the total amount of data collected and is crucial for evaluating the performance of the accelerator. Integrated luminosity is defined as~\cite{concept_of_luminosity}:

\begin{equation}
  \mathcal{L}_{\text{int}} = \int_{0}^{T} \mathcal{L}_{\text{inst}}(t')\,dt'
\end{equation}

Here, $\mathcal{L}_{\text{inst}}$ is evaluated over the duration $T$ of the data-taking period (excluding possible dead time), as its intensity typically decreases over time due to the reduction in the number of protons during the fill.

@@

\section{Current Luminosity precision for LHC Run 2}

@@

The second stage (2017–2018) had preliminary results published in the Public Analysis Summaries for 2017 and 2018 \cite{pas_17,pas_18}. However, the desired uncertainty values were not achieved.
Efforts are currently underway to further improve the precision of these measurements and present them in a final publication. This project is primarily focused on contributing to this publication by analyzing the PCC luminometer and refining the uncertainty, aiming to achieve values of approximately 1\%.\\

@@

### 2025-05-06 monograph_Luis/Portada/Portada.tex

 %% Aquí comienza el documento.
\thispagestyle{empty} %% PARA QUITAR FORMATO DE PIES Y ENCABEZADOS DE PÁGINA A LA PORTADA
%
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO IZQUIERDO (UNIVERSIDAD) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[c][0.15\textheight][c]{0.25\textwidth}
    \begin{center}
        \includegraphics[height=3.2cm, keepaspectratio=true]{unison-logo.png} %% LOGO IZQUIERDO
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%% MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TITULOS, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%
%
%
%
%\fbox{
\begin{minipage}[c][0.17\textheight][t]{0.7\textwidth}
    \begin{center}
        {\scshape\LARGE UNIVERSIDAD DE SONORA} %% NOMBRE DE LA UNIVERSIDAD
        \vspace{.5cm}   %% DEFINE EL ESPACIO ENTRE EL NOMBRE DE LA UNIVERSIDAD Y LA LÍNEA DOBLE
        \hrule height2.5pt  %% DEFINE EL ANCHO DE LA PRIMER LÍNEA
        \vspace{.1cm}    %% DEFINE EL ESPACIO ENTRE LAS DOS LINEAS
        \hrule height1pt  %% DEFINE EL ANCHO DE LA SEGUNDA LINEA (NORMALMENTE ES MAS DELGADA)
        \vspace{.4cm}   %% DEFINE EL ESPACIO ENTRE LA DOBLE LINEA Y EL NOMBRE DE LA DIVISIÓN
        {\scshape \large Facultad Interdisciplinaria de Ciencias Exactas y Naturales}
        {\scshape Departamento de Investigaci\'on en F\'isica} %% NOMBRE DE LA DIVISIÓN
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO DERECHO (DIVISIÓN ACADÉMICA)
%
%
%
%
%\fbox{

@@

        \makebox[5cm][c]{\LARGE Monograf\'ia}  \\[20pt]
         En cumplimiento parcial de los requisitos para el grado de:\\[5pt]
        {\Large \textbf{{Doctorado en Ciencias (F\'isica)}}}\\[40pt]
        By:\\[12pt]
        \textbf{  {\large Luis Enriquue Enrique CUEVAS PICOS}}

>>>>

 %% Aquí comienza el documento.
\thispagestyle{empty} %% PARA QUITAR FORMATO DE PIES Y ENCABEZADOS DE PÁGINA A LA PORTADA
%
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO IZQUIERDO (UNIVERSIDAD) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[c][0.15\textheight][c]{0.25\textwidth}
    \begin{center}
        \includegraphics[height=3.2cm, keepaspectratio=true]{unison-logo.png} %% LOGO IZQUIERDO
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%% MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TITULOS, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%
%
%
%
%\fbox{
\begin{minipage}[c][0.17\textheight][t]{0.7\textwidth}
    \begin{center}
        {\scshape\LARGE UNIVERSIDAD DE SONORA} %% NOMBRE DE LA UNIVERSIDAD
        \vspace{.5cm}   %% DEFINE EL ESPACIO ENTRE EL NOMBRE DE LA UNIVERSIDAD Y LA LÍNEA DOBLE
        \hrule height2.5pt  %% DEFINE EL ANCHO DE LA PRIMER LÍNEA
        \vspace{.1cm}    %% DEFINE EL ESPACIO ENTRE LAS DOS LINEAS
        \hrule height1pt  %% DEFINE EL ANCHO DE LA SEGUNDA LINEA (NORMALMENTE ES MAS DELGADA)
        \vspace{.4cm}   %% DEFINE EL ESPACIO ENTRE LA DOBLE LINEA Y EL NOMBRE DE LA DIVISIÓN
        {\scshape \large Facultad Interdisciplinaria de Ciencias Exactas y Naturales}\\
        {\scshape Departamento de Investigaci\'on en F\'isica} %% NOMBRE DE LA DIVISIÓN
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO DERECHO (DIVISIÓN ACADÉMICA)
%
%
%
%
%\fbox{

@@

        \makebox[5cm][c]{\LARGE Monograf\'ia}  \\[20pt]
         En cumplimiento parcial de los requisitos para el grado de:\\[5pt]
        {\Large \textbf{{Doctorado en Ciencias (F\'isica)}}}\\[40pt]
        By:\\[12pt]
        \textbf{  {\large Luis Enrique CUEVAS PICOS}}
